---
title: "Rock climbing survey data claning & analysis"
author: "Sunny Tseng"
date: "2023-03-05"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data description

There are 3 cohorts experienced rock climbing intervention. Surveys were done before intervention, after intervention, and 1-year after intervention (only for cohort 1). This research is aim to understand how the intervention would influence the behaviour of the participants. There were 20 questions asked, with 4 different levels: Strongly Agree, Agree, Disagree, Strongly Disagree. The goal of the analysis is to test whether there are significant differences between those levels before and after the intervention. 


### R packages

Here we install (by `install.packages()`) and load (by `library()`) the packages that we will need. A package is like a tool box in R, that includes many functions. For example, `tidyverse` package provides functions for data wrangling and data cleaning, `here` packages provides functions for working directory management.  

```{r, message = FALSE}
library(tidyverse)
library(readxl)
library(here)
library(RColorBrewer)
library(coin)
```


## Cohort 1

### Import data

Use `read_excel` to import raw data. Note that the name of the folder and the file name were changed as there were parentheses `()` and commas `,` in the original name, which causes error when importing files into R. 

```{r, message = FALSE}
data_c1_s1 <- read_excel(here("data", 
                              "Cohort 1 Surveys_123_spreadsheets", 
                              "VIMFF Cohort 1 Survey 1_pre-intervention.xlsx"))

data_c1_s2 <- read_excel(here("data", 
                              "Cohort 1 Surveys_123_spreadsheets", 
                              "VIMFF Cohort 1 Survey 2_post-intervention.xlsx"))

data_c1_s3 <- read_excel(here("data", 
                              "Cohort 1 Surveys_123_spreadsheets", 
                              "VIMFF Cohort 1 Survey 3_post-post.xlsx"))

```


### Data cleaning

Cleaned column names, combined columns with multiple choice answers, changed the categorical answer into 1 - 4 for easier analysis. Only select columns that will need in the following analysis. 

For cohort 1, survey 1:
```{r}
new_name <- function(dataset, start, end){
  dataset[1, start:end] %>% as_vector()
}

data_c1_s1_clean <- data_c1_s1 %>%
  rename_with(.fn =~ new_name(data_c1_s1, 65, 84), .cols = c(65:84)) %>%
  filter(row_number() != 1) %>%
  unite(gender, names(.)[11:23], sep = ",", na.rm = TRUE) %>%
  unite(nationality, names(.)[12:28], sep = ",", na.rm = TRUE) %>%
  unite(indoor_style_practice, names(.)[24:26], sep = ",", na.rm = TRUE) %>%
  unite(outdoor_style_practice, names(.)[29:34], sep = ",", na.rm = TRUE) %>%
  mutate(cohort = "c1", survey = "s1") 

data_c1_s1_clean_1 <- data_c1_s1_clean %>%
  select("Respondent ID", "Collector ID", "indoor_style_practice",
         "outdoor_style_practice", 30:49, "cohort", "survey") 

#data_c1_s1_clean %>% str()
```

For cohort 1, survey 2:
```{r}
data_c1_s2_clean <- data_c1_s2 %>%
  rename_with(.fn =~ new_name(data_c1_s2, 32, 51), .cols = c(32:51)) %>%
  filter(row_number() != 1) %>%
  unite(indoor_style_practice, names(.)[19:21], sep = ",", na.rm = TRUE) %>%
  unite(outdoor_style_practice, names(.)[24:29], sep = ",", na.rm = TRUE) %>%
  mutate(cohort = "c1", survey = "s2") %>%
  select("Respondent ID", "Collector ID", "indoor_style_practice",
         "outdoor_style_practice", 25:44, "cohort", "survey") 

#data_c1_s2_clean %>% str()
```

For cohort 1, survey 3:
```{r}
data_c1_s3_clean <- data_c1_s3 %>%
  rename_with(.fn =~ new_name(data_c1_s3, 32, 51), .cols = c(32:51)) %>%
  filter(row_number() != 1) %>%
  unite(indoor_style_practice, names(.)[19:21], sep = ",", na.rm = TRUE) %>%
  unite(outdoor_style_practice, names(.)[24:29], sep = ",", na.rm = TRUE) %>%
  mutate(cohort = "c1", survey = "s3") %>%
  select("Respondent ID", "Collector ID", "indoor_style_practice",
         "outdoor_style_practice", 25:44, "cohort", "survey") 

#data_c1_s3_clean %>% str()
```

For cohort 1 (final summary). Note that there are some mismatch between column names. Need to fix that before combining data frames. Here we only use before & after interventation for demonstration. We also compiled a dataframe to show individual participant information. 

```{r}
# s1 says "I find it difficult to find transportation to go outdoor climbing", while
# s2 says "I find it difficult to find transportation to outdoor climbing"
# There are many mis-match between s3 and s1&s2. Need check!

names(data_c1_s1_clean_1)[23] <- names(data_c1_s2_clean)[23] 
data_c1_clean <- rbind(data_c1_s1_clean_1, data_c1_s2_clean) %>%
    mutate_at(c(5:24), ~recode(., "Strongly Agree" = 4,
                             "Strongly agree" = 4,
                             "Agree" = 3,
                             "Disagree" = 2,
                             "Strongly Disagree" = 1,
                             "Strongly disagree" = 1),
            na.rm = TRUE) %>%
    mutate_at(c(5:24), ~recode(., "4" = "Strongly Agree",
                               "3" = "Agree",
                               "2" = "Disagree",
                               "1" = "Strongly Disagree")) %>%
    mutate_at(c(5:24), ~ factor(., levels = c("Strongly Disagree",
                                              "Disagree",
                                              "Agree",
                                              "Strongly Agree"))) 


data_c1_clean %>% str()
```

```{r}
data_c1_clean_individual <- data_c1_s1_clean %>%
  select("Respondent ID", "gender", "nationality", 13:18)

data_c1_clean_individual %>% str()
```


### Data visualization

Positive questions for climbing - before & after intervention

```{r}
nb.cols <- 20
mycolors <- colorRampPalette(brewer.pal(8, "Set2"))(nb.cols)

positive_c1_fig <- data_c1_clean %>%
  filter(cohort == "c1") %>%
  select(`Respondent ID`, survey, 5:8, 10:20) %>%
  pivot_longer(!c(`Respondent ID`, survey), 
               names_to = "question", 
               values_to = "Level of agreement") %>%
  mutate(survey = if_else(survey == "s1", "Before intervention", "After intervention")) %>%
  mutate(survey = factor(survey, levels = c("Before intervention", "After intervention"))) %>%
  drop_na() %>%
  ggplot() +
    geom_bar(aes(`Level of agreement`, fill = question)) + # position = position_dodge()
    scale_fill_manual(values = mycolors) +
    facet_grid(~survey) +
    theme(legend.position = "none") 

positive_c1_fig

```

```{r}
positive_c1_table <- data_c1_clean %>%
  filter(cohort == "c1") %>%
  select(`Respondent ID`, survey, 5:8, 10:20) %>%
  pivot_longer(!c(`Respondent ID`, survey), 
               names_to = "question", 
               values_to = "agreement") %>%
  drop_na() %>%
  group_by(agreement, survey) %>%
  summarize(n = n()) %>%
  pivot_wider(names_from = survey, values_from = n) %>%
  mutate(percent_before = s1/sum(.$s1)*100,
         percent_after = s2/sum(.$s2)*100) %>%
  rename(n_before = "s1", n_after = "s2") %>%
  select(agreement, n_before, percent_before, n_after, percent_after)

positive_c1_table
```


Negative questions for climbing - before & after intervention

```{r}
nb.cols <- 20
mycolors <- colorRampPalette(brewer.pal(8, "Set2"))(nb.cols)

negative_c1 <- data_c1_clean %>%
  filter(cohort == "c1") %>%
  select(`Respondent ID`, survey, 9, 21:24) %>%
  pivot_longer(!c(`Respondent ID`, survey), 
               names_to = "question", 
               values_to = "Level of agreement") %>%
  mutate(survey = if_else(survey == "s1", "Before intervention", "After intervention")) %>%
  mutate(survey = factor(survey, levels = c("Before intervention", "After intervention"))) %>%
  drop_na() %>%
  ggplot() +
    geom_bar(aes(`Level of agreement`, fill = question)) + # position = position_dodge()
    scale_fill_manual(values = mycolors) +
    facet_grid(~survey) +
    theme(legend.position = "none") 

negative_c1
```

```{r}
negative_c1_table <- data_c1_clean %>%
  filter(cohort == "c1") %>%
  select(`Respondent ID`, survey, 9, 21:24) %>%
  pivot_longer(!c(`Respondent ID`, survey), 
               names_to = "question", 
               values_to = "agreement") %>%
  drop_na() %>%
  group_by(agreement, survey) %>%
  summarize(n = n()) %>%
  pivot_wider(names_from = survey, values_from = n) %>%
  mutate(percent_before = s1/sum(.$s1)*100,
         percent_after = s2/sum(.$s2)*100) %>%
  rename(n_before = "s1", n_after = "s2") %>%
  select(agreement, n_before, percent_before, n_after, percent_after)

negative_c1_table
```

### Statistical test for paired ordinal data

#### Some good online resources:

- [What types of statistical test can be used for paired categorical variables](https://www.researchgate.net/post/What-types-of-statistical-test-can-be-used-for-paired-categorical-variables-For-more-than-two-category)

- [What is the most suitable statistical test for ordinal data](https://www.researchgate.net/post/What_is_the_most_suitable_statistical_test_for_ordinal_data_eg_Likert_scales#:~:text=The%20most%20suitable%20statistical%20tests,%2C%20no%20assumption%20on%20distribution)

- [Choosing the Right Statistical Test](https://www.scribbr.com/statistics/statistical-tests/#:~:text=They%20can%20be%20used%20to%20test%20the%20effect%20of%20a,heights%20of%20men%20and%20women)

- [Can You Use a T-Test on Ranked Data?](https://sciencing.com/can-use-ttest-ranked-data-12010046.html)

#### T-test

>T-tests are not appropriate to use with ordinal data. Because ordinal data has no central tendency, it also has no normal distribution. The values of ordinal data are evenly distributed, not grouped around a mid-point. Because of this, a t-test of ordinal data would have no statistical meaning.

#### Wilcoxon Signed Rank Test

>The Wilcoxon signed rank test (also called the Wilcoxon signed rank sum test) is a non-parametric test to compare data. The Wilcoxon signed rank test should be used if the differences between pairs of data are non-normally distributed. The Wilcoxon matched-pairs signed rank test computes the difference between each set of matched pairs, then follows the same procedure as the signed rank test to compare the sample against some median. The null hypothesis for this test is that the medians of two samples are equal.



### Wilcoxon matched-pairs signed rank test for single question

The following analysis is based on the instructions provided by [Ordinal vs Ordinal paired](https://peterstatistics.com/CrashCourse/4-TwoVarPair/OrdOrd/OrdOrdPair0.html). Here we demonstrate a single question, say the 5th column. 

First we summarize the response of the participants:
```{r}
wilcox_1 <- data_c1_clean %>%
  select(`Respondent ID`, 5, `survey`) %>%
  mutate_at(c(2), ~recode(., "Strongly Agree" = 4,
                             "Agree" = 3,
                             "Disagree" = 2,
                             "Strongly Disagree" = 1)) %>%
  pivot_wider(names_from = survey, values_from = `I love to climb`) 

wilcox_1
```


Then we plot out the bar plot of difference to see the trend of change between before and after. 
```{r, warning = FALSE}
wilcox_1_box <- wilcox_1 %>%
  mutate(difference = s2 - s1) %>%
  ggplot() +
    geom_bar(aes(difference))

wilcox_1_box
```

Finally we conduct a statistical test and look for p-value, with the null hypothesis being **the medians of two samples are equal**.
```{r, warning = FALSE}
wilcox.test(x = wilcox_1$s1, y = wilcox_1$s2,
            paired = TRUE,
            conf.int = TRUE,
            conf.level = 0.95)
```


```{r}
wilcoxsign_test(wilcox_1$s1 ~ wilcox_1$s2, zero.method = "Wilcoxon")
```

### Questions for the dataset :)

- Are the data of the 3 cohorts need to be pooled? Or each of the cohort need to be compared seperately? 
- There are some NAs in the reply? 
- Pool the questions, or by group (positive & negative), or by single question?







